% !TEX root = ../main.tex
%-------------------------------------------------------------------------------
\subsection{Computational implementation}\label{Computational implementation}
%-------------------------------------------------------------------------------
We use the same computational implementation as in \citet{Keane.1997}. We outline the per-period utility functions for each of the five alternatives. We first focus on their common overall structure and then provide their parameterization. Throughout we provide the economic motivation for their specification.\\

\noindent Individuals are initially heterogeneous. They differ with respect to their initial level of completed schooling $h_1$ and have one of four different $\mathcal{J} = \{1, \hdots, 4\}$ alternative-specific skill endowments $\bm{e} = \left(e_{j,a}\right)_{\mathcal{J} \times \mathcal{A}}$. However, individuals share the same discounting parameter $\theta$. \\

\noindent The per-period utility $u_a(\cdot)$ of each alternative consists of a non-pecuniary utility $\zeta_a(\cdot)$ and, at least for the working alternatives, an additional wage component $w_a(\cdot)$. Each of the components depends on the level of human capital as measured by their alternative-specific skill endowment $\bm{e}$, years of completed schooling $h_t$, and occupation-specific work experience $\bm{k_t} = \left(k_{a,t}\right)_{a\in\{1, 2, 3\}}$. The per-period utility functions are also influenced by last-period choices $a_{t -1}$ and alternative-specific productivity shocks $\bm{\epsilon_t} = \left(\epsilon_{a,t}\right)_{a\in\mathcal{A}}$. Throughout alternatives the general form of the per-period utility is given by:
%
\begin{align*}
u_a(\cdot) =
\begin{cases}
    \zeta_a(\bm{k_t}, h_t, t, a_{t -1})  + w_a(\bm{k_t}, h_t, t, a_{t -1}, e_{j, a}, \epsilon_{a,t})                & \text{if}\, a \in \{1, 2, 3\}  \\
    \zeta_a(\bm{k_t}, h_t, t, a_{t-1}, e_{j,a}, \epsilon_{a,t})                                                  &  \text{if}\, a \in \{4, 5\}.
\end{cases}
\end{align*}
%
Work experience $\bm{k_t}$  and years of completed schooling $h_t$ evolve deterministically.
%
\begin{align*}
k_{a,t+1} = k_{a,t} + \ind[a_t = a]  &\qquad \text{if}\, a \in \{1, 2, 3\} \\
h_{t + 1\phantom{,a}} = h_{t\phantom{,a}} +   \ind[a_t = 4]  &\qquad
\end{align*}
%-------------------------------------------------------------------------------
\subsubsection*{Per-period utility for working alternatives}
%-------------------------------------------------------------------------------
The per-period utility for working alternatives is the sum of a non-pecuniary utility and a wage component. The wage component $w_{a}(\cdot)$ is given by the product of the market-equilibrium rental price $r_{a}$ and an occupation-specific skill level $x_{a}(\cdot)$. The overall level of human capital determines the latter.
%
\begin{align*}
w_{a}(\cdot) = r_{a} \, x_{a}(\cdot)
\end{align*}
%
This specification leads to a standard logarithmic wage equation in which the constant term is the skill rental price $\ln(r_{a})$.\\

\noindent The occupation-specific skill level $x_{a}(\cdot)$ is determined by a skill production function, which includes a deterministic component $\Gamma_a$ and a multiplicative stochastic productivity shock $\epsilon_{a,t}$.
%
\begin{align}%\label{eq:OccupationSpecificSkillLevel}
    x_{a}(\bm{k_t}, h_t, t, a_{t-1}, e_{j, a}, \epsilon_{a,t}) & = \exp \big( \Gamma_{a}(\bm{k_t},  h_t, t, a_{t-1}, e_{j,a}) \cdot \epsilon_{a,t} \big) \nonumber
\end{align}
%
\noindent\textbf{White-collar} Equation (\ref{eq:SkillLevelWhiteCollar}) shows the parameterization of the deterministic component of the skill-production function.
%
\begin{align}\label{eq:SkillLevelWhiteCollar}
    \Gamma_1(\bm{k_t}, h_t, t, a_{t-1}, e_{j, 1}) = e_{j,1} & + \beta_{1,1} \cdot h_t + \beta_{1, 2} \cdot \ind[h_t \geq 12] + \beta_{1,3} \cdot \ind[h_t\geq 16]\\ \nonumber
                                  & + \gamma_{1, 1} \cdot  k_{1,t} + \gamma_{1,2} \cdot  (k_{1,t})^2 + \gamma_{1,3} \cdot  \ind[k_{1,t} > 0] \\ \nonumber
                                & + \gamma_{1,4} \cdot  t + \gamma_{1,5} \cdot \ind[t < 18]\\ \nonumber
                                  & + \gamma_{1,6} \cdot \ind[a_{t-1} = 1] + \gamma_{1,7} \cdot  k_{2,t} + \gamma_{1,8} \cdot  k_{3,t} \nonumber
\end{align}
%
\noindent Part of the skill-production function is motivated by \citet{Mincer.1958, Mincer.1974} and hence linear in years of completed schooling $\beta_{1,1}$, quadratic in experience ($\gamma_{1,1}, \gamma_{1,2}$), and separable between the two of them. The skill production-function respects sheep-skin effects ($\beta_{1,2}, \beta_{1,3}$) and the gain of having worked in the same occupation at least once before $\gamma_{1,3}$ \citep{Spence.1973, Jaeger.1996}. Skill production depends linearly on the age $\gamma_{1,4}$ and on being a minor $\gamma_{1,5}$. Work experience depreciates when working in a different occupation ($\gamma_{1,6}$) but is potentially transferable across occupations ($\gamma_{1,7}, \gamma_{1,8}$). \\


\noindent Equation (\ref{eq:NonWageWhiteCollar}) shows the parameterization of the non-pecuniary utility from working in a white-collar occupation.
%
\begin{align}\label{eq:NonWageWhiteCollar}
\zeta_{1}(\bm{k_t}, h_t, a_{t-1})  = \alpha_1  &+ c_{1,1} \cdot \ind[a_{t-1} \neq 1] + c_{1,2} \cdot \ind[k_{1,t} = 0] \\ \nonumber
                            & + \vartheta_1 \cdot \ind[h_t \geq 12] + \vartheta_2 \cdot \ind[h_t \geq 16] + \vartheta_3 \cdot \ind[k_{3,t} = 1]
\end{align}
%
A constant $\alpha_1$ captures the net monetary-equivalent of on the job amenities. The non-pecuniary utility includes mobility and search costs $c_{1,1}$, which are higher for individuals who never worked in a white-collar occupation before $c_{1,2}$. The non-pecuniary components capture returns from a high school degree $\vartheta_1$ and a college degree $\vartheta_2$. Additionally, there is a detrimental effect of leaving the military early, i.e., after one year $\vartheta_3$.\\

\noindent\textbf{Blue-collar} The per-period utility from working in a blue-collar occupation is specified analogously. Equation (\ref{eq:SkillLevelBlueCollar}) shows the parameterization of the deterministic component of the skill-production function.
%
\begin{align}\label{eq:SkillLevelBlueCollar}
    \Gamma_2(\bm{k_t}, h_t, t, a_{t-1}, e_{j,2}) = e_{j,2} & + \beta_{2,1} \cdot h_t + \beta_{2, 2} \cdot \ind[h_t \geq 12] + \beta_{2,3} \cdot \ind[h_t\geq 16] \\\nonumber
    							 & + \gamma_{2, 1} \cdot  k_{2,t} + \gamma_{2,2} \cdot  (k_{2,t})^2 + \gamma_{2,3} \cdot  \ind[k_{2,t} > 0] \\\nonumber
                                   & + \gamma_{2,4} \cdot  t + \gamma_{2,5} \cdot \ind[t < 18] \\\nonumber
                                  & + \gamma_{2,6} \cdot  \ind[a_{t-1} = 2]  + \gamma_{2,7} \cdot  k_{1,t} + \gamma_{2,8} \cdot  k_{3,t}.
\end{align}
%
\noindent Equation (\ref{eq:UtilityBlueCollar}) shows the parameterization of the non-pecuniary utilities from working in a blue-collar occupation.
%
\begin{align}\label{eq:UtilityBlueCollar}
\zeta_{2}( \bm{k_t}, h_t, a_{t-1} ) = \,\alpha_2 & + c_{2,1} \cdot \ind[a_{t-1} \neq 2] + c_{2,2} \cdot \ind[k_{2,t} = 0]\\\nonumber
                            & + \vartheta_1 \cdot \ind[h_t \geq 12] + \vartheta_2 \cdot \ind[h_t \geq 16] + \vartheta_3 \cdot \ind[k_{3,t} = 1].
\end{align}
%
\noindent\textbf{Military} Equation (\ref{eq:SkillLevelMilitary}) shows the parameterization of the deterministic component of the skill-production function.
%
\begin{align}\label{eq:SkillLevelMilitary}
    \Gamma_3( k_{3,t}, h_t, t, e_{j,3}) = e_{j,3} & + \beta_{3,1} \cdot h_t \\\nonumber
	               \nonumber &+ \gamma_{3,1} \cdot  k_{3,t} + \gamma_{3,2} \cdot (k_{3,t})^2 + \gamma_{3,3} \cdot \ind[k_{3,t} > 0]\\\nonumber
									 & + \gamma_{3,4} \cdot t + \gamma_{3,5} \cdot \ind[t < 18].
\end{align}
%
Contrary to the civilian sector there are no sheep-skin effects from graduation ($\beta_{3,2} = \beta_{3,3}= 0$). The previous occupation choice has no influence ($\gamma_{3,6}= 0$) and any experience other than military is non-transferable ($\gamma_{3,7} = \gamma_{3,7} = 0$).\\

\noindent Equation (\ref{eq:UtilityMilitary}) shows the parameterization of the non-pecuniary utility from working in the military.
%
\begin{align}\label{eq:UtilityMilitary}
\zeta_{3}( k_{3.t}, h_t)  = \,& c_{3,2} \cdot \ind[k_{3,t} = 0]+ \vartheta_1 \cdot \ind[h_t \geq 12] + \vartheta_2 \cdot \ind[h_t \geq 16]
\end{align}
%
Search costs $(c_{1, 1} = c_{1, 2} = 0)$ are absent but it is costly if an individual has never served in the military before $c_{3,2}$. Individuals still experience a non-pecuniary utility finishing high-school $\vartheta_1$ and college $\vartheta_2$.


\begin{Remark} Our parameterization for the per-period utility of serving in the military differs from \citet{Keane.1997} as we remain unsure about their exact specification. The authors state in Footnote 31 (p.498) that the constant for the non-pecuniary utility $\alpha_{3,t}$ depends on age. However, we are unable to determine the precise nature of the relationship. Also, Equation (C3) (p.521) indicates no productivity shock $\epsilon_{a,t}$ in the wage component. However, Table 7 (p.500) reports such estimates.
\end{Remark}
% ------------------------------------------------------------------------------
\FloatBarrier\subsubsection*{Per-period utility for non-working alternatives}
% ------------------------------------------------------------------------------
The per-period utility for the non-working alternatives does not contain a wage component. Also, the productivity shock enters in an additive instead of a multiplicative fashion.
%-------------------------------------------------------------------------------
\FloatBarrier\paragraph*{School}
%-------------------------------------------------------------------------------
Equation (\ref{eq:UtilitySchooling}) shows the parameterization of the non-pecuniary utility from schooling.
%
\begin{align}\label{eq:UtilitySchooling}
	\zeta_4(k_{3,t}, h_t, t, a_{t-1}, e_{j,4}, \epsilon_{4,t})  = e_{j,4} & + \beta_{tc_1} \cdot \ind[h_t \geq 12] + \beta_{tc_2} \cdot \ind[h_t \geq 16]   \\\nonumber
    							  & + \beta_{rc_1} \cdot \ind[a_{t-1} \neq 4, h_t < 12] + \beta_{rc_2} \cdot \ind[a_{t-1} \neq 4, h_t \geq 12] \\\nonumber
    							  & + \gamma_{4,4} \cdot t + \gamma_{4,5} \cdot \ind[t < 18] 																					  \\\nonumber
     							  & + \vartheta_1 \cdot \ind[h_t \geq 12] + \vartheta_2 \cdot \ind[h_t \geq 16] + \vartheta_3 \cdot \ind[k_{3,t} = 1)\\\nonumber
      							& + \epsilon_{4,t}
\end{align}
%
Schooling costs include tuition and fees for continuing education after high school $(\beta_{tc_1}, \beta_{tc_2})$. Leaving the schooling alternative is reversible but returning entails adjustment cost that differs by educational category ($\beta_{rc_1}, \beta_{rc_2}$). Although schooling is defined as time spent in school, individuals value completion of high-school and graduate school ($\vartheta_1, \vartheta_2$). Once individuals reach a certain amount of schooling, they obtain a grade, i.e., there is no uncertainty about grade completion \citep{Altonji.1993}. Choices are mutually exclusive, and so there is no part-time enrollment.
%-------------------------------------------------------------------------------
\FloatBarrier\paragraph*{Home}
%-------------------------------------------------------------------------------
Equation (\ref{eq:UtilityHome}) shows the parameterization of the non-pecuniary utility from staying at home.
%
\begin{align}\label{eq:UtilityHome}
	\zeta_5(k_{3,t}, h_t, t, e_{j,5}, \epsilon_{5,1}) =  e_{j,5} & + \gamma_{5,4} \cdot \ind[18 \leq t \leq 20] + \gamma_{5,5} \cdot \ind[t \geq 21] \\ \nonumber
    							   & +\vartheta_{1} \cdot \ind[h_t \geq 12] + \vartheta_{2} \cdot \ind[h_t \geq 16] +  \vartheta_3 \cdot \ind[k_{3,t} = 1]  \\ \nonumber
    							   & + \epsilon_{5,t}.
\end{align}
%
Staying at home as a young adult $\gamma_{5, 4}$ is perceived to be less stigmatic as doing so while being already an adult ($\gamma_{5,5}$). Having a degree  $(\vartheta_1, \vartheta_2)$ or having left the military prematurely  $\vartheta_3$ does influence the per-period utility.

\paragraph{Distribution of unobservables} The productivity shocks are uncorrelated across time and follow a multivariate normal distribution $\mathcal{N}$ with mean $\bm{0}$ and covariance matrix $\bm{\Sigma}$.

\paragraph{State space} Given the structure of the utility functions and the distribution of the shocks, the state at time $t$ is $s_t = \{\bm{k_t}, h_t, t, a_{t -1}, \bm{e},\bm{\epsilon_t}\}$.

%-------------------------------------------------------------------------------
\FloatBarrier\subsubsection*{Overview parameters}
%-------------------------------------------------------------------------------

\input{material/tab-model-parameters}
